---
title: "4355 Project"
author: "Alondra Ramos"
output:
  word_document: default
  html_document:
    df_print: paged
---
```{r, packages}
library(tidyverse)
library(reshape2)
```

# Data Cleaning
```{r, data cleaning}
df <- read.csv("https://raw.githubusercontent.com/alondra170/stat-4355-project/main/Top500.csv")
#rename title
colnames(df)[1] <- gsub('^...','',colnames(df)[1])
#there are some duplicate rows, we need to remove them
df <- df[!duplicated(df), ]
#change dur var from char to numeric
df$dur <- as.numeric(gsub(",","",df$dur))
#remove added column; useless
df <- as_tibble(df) %>% dplyr::select(-added)
summary(df)
glimpse(df)
df %>% filter(bpm < 10)
```

# data exploration
```{r, data exploration}
ggplot(df, aes(x = nrgy)) +
  geom_histogram(aes(y = stat(count / sum(count))), fill = "#009E73", bins = 100) +
  labs(title="Song Energy Histogram", y="Frequency", x="Energy Level") + geom_density()
```

```{r}
numlabs <- c("year", "bpm", "nrgy", "dnce", "dB", "live", "val", "dur", "acous", "spch", "pop")

plots <- function(var) {
labs <- numlabs[which(numlabs != var)]
dfmelty <- df %>% melt(measure.vars = labs, id.vars = var)
ggplot(dfmelty, aes_string(x = "value", y = var)) + geom_point(shape = ".") + facet_wrap(~variable, scales="free_x") + geom_smooth(method = lm)
}
lapply(numlabs, plots)
```

```{r}
ggplot(dfmelty) +
geom_jitter(aes(value,live, colour=variable),) +
geom_smooth(aes(value,live, colour=variable), method=lm, se=FALSE) +
facet_wrap(~variable, scales="free_x")

dfmelty <- melt(df.1[,1:10], id.vars = 'nrgy')

ggplot(dfmelty) +
geom_jitter(aes(value,nrgy, colour=variable),) +
geom_smooth(aes(value,nrgy, colour=variable), method=lm, se=FALSE) +
facet_wrap(~variable, scales="free_x")

dfmelty <- melt(df.1[,1:10], id.vars = 'acous')

ggplot(dfmelty) +
geom_jitter(aes(value,acous, colour=variable),) +
geom_smooth(aes(value,acous, colour=variable), method=lm, se=FALSE) +
facet_wrap(~variable, scales="free_x")

dfmelty <- melt(df.1[,1:10], id.vars = 'spch')

ggplot(dfmelty) +
geom_jitter(aes(value,spch, colour=variable),) +
geom_smooth(aes(value,spch, colour=variable), method=lm, se=FALSE) +
facet_wrap(~variable, scales="free_x")

dfmelty <- melt(df.1[,1:10], id.vars = 'dB')

ggplot(dfmelty) +
geom_jitter(aes(value,dB, colour=variable),) +
geom_smooth(aes(value,dB, colour=variable), method=lm, se=FALSE) +
facet_wrap(~variable, scales="free_x")

dfmelty <- melt(df.1[,1:10], id.vars = 'bpm')

ggplot(dfmelty) +
geom_jitter(aes(value,bpm, colour=variable),) +
geom_smooth(aes(value,bpm, colour=variable), method=lm, se=FALSE) +
facet_wrap(~variable, scales="free_x")

dfmelty <- melt(df.1[,1:10], id.vars = 'year')

ggplot(dfmelty) +
geom_jitter(aes(value, year, colour=variable),) +
geom_smooth(aes(value, year, colour=variable), method=lm, se=FALSE) +
facet_wrap(~variable, scales="free_x")

dfmelty <- melt(df.1[,1:10], id.vars = 'dnce')

ggplot(dfmelty) +
geom_jitter(aes(value, dnce, colour=variable),) +
geom_smooth(aes(value, dnce, colour=variable), method=lm, se=FALSE) +
facet_wrap(~variable, scales="free_x")

dfmelty <- melt(df.1[,1:10], id.vars = 'val')

ggplot(dfmelty) +
geom_jitter(aes(value, val, colour=variable),) +
geom_smooth(aes(value, val, colour=variable), method=lm, se=FALSE) +
facet_wrap(~variable, scales="free_x")
```



# initial fitting
```{r}
fit.1 <- lm(nrgy~., data = df.1)
summary(fit.1)

```

# feature selection
```{r}
library(car)
vif(fit.1)
```

```{r}
cor(df.1[,c(1:2, 4:10)])
```
```{r}
par(mfrow = c(2, 2))
plot(fit.1)
```
# residual analysis
```{r}
library(MASS)

barplot(height = stdres(fit.1), names.arg = 1:6888,
        main = "Standardized Residuals", xlab = "Index",
        ylab = "Standardized Resid", ylim = c(-4, 4))

# cutoff values
abline(h = 3, col = "Red", lwd = 3)
abline(h = -3, col = "Red", lwd = 3)
```
# studentized residuals
```{r}
barplot(height = studres(fit.1), names.arg = 1:6888,
        main = "Studentized Residuals", xlab = "Index",
        ylab = "Studentized Resid", ylim = c(-4,4))

abline(h = 3, col = "Red", lwd = 3)
abline(h = -3, col = "Red", lwd = 3)
```
# r student residuals
```{r}
barplot(height = rstudent(fit.1), names.arg = 1:6888,
        main = "R Student Residuals", xlab = "Index",
        ylab = "R Student Resid", ylim = c(-5, 5))

cor.level <- 0.05/(2 * 6888)
cor.qt <- qt(cor.level, 6878, lower.tail=F)

abline(h = cor.qt, col = "Red", lwd = 3)
abline(h = -cor.qt, col = "Red", lwd = 3)
```
## Influential Analysis
```{r}
myInf <- influence.measures(fit.1)
summary(myInf)
```
```{r}
library(car)
dfbetasPlots(fit.1, intercept = F)
```
```{r}
influenceIndexPlot(fit.1)
```
## Fittings with suspected outliers removed

```{r}
# removing suspected outliers
df.red.1 <- df.1[-c(6671, 6685, 6686),]

fit.2 <- lm(nrgy~., data = df.red.1)
summary(fit.2)
```
```{r}
par(mfrow = c(2, 2))
plot(fit.2)
```
##  Testing for heteroscedascity
```{r}
library(lmtest)
bptest(fit.2)

```


## Testing Transformations

Although there are some improvements in the R^2 values with the transformations, there seems to be a loss in the results for the residuals that indicate a slightly poorer fit.

# log(nrgy)

```{r}
fit.3 <- lm(log(nrgy)~., data = df.red.1) #not knitting for me
summary(fit.3)
```

```{r}
par(mfrow = c(2, 2))
plot(fit.3)
```
# sqrt(nrgy)
```{r}
fit.4 <- lm(sqrt(nrgy)~., data = df.red.1)
summary(fit.4)
```
```{r}
par(mfrow = c(2, 2))
plot(fit.3)
```
```{r}
wt <- 1 / lm(abs(fit.2$residuals) ~ fit.2$fitted.values)$fitted.values^2
fit.wls <- lm(nrgy~., data = df.red.1, weights = wt)
summary(fit.wls)
```


